createData_linear_modeling

### 2016-06-14 portfolio/analysis/1.5-createData_linear_modeling.Rmd
Start rearranging the files

cfecAnnual = group_by(cfecnew, p_holder, year) %>%
  summarize(permit = p_fshy[1], boat = cadfg2[1], 
    days = diff(range(day) + 1), 
    ngear = length(unique(gearn)),
    nspec = length(unique(specn)),
    npermit = length(unique(p_fshy)),
    revenue = sum(g_earn),
    length = mean(VLength, na.rm=T)) %>% 
  filter(npermit == 1)

@@

temp = group_by(cfecnew, p_holder, year, specn) %>% 
  summarize(totRev.spec = sum(g_earn), 
    permit = p_fshy[1])
temp = left_join(temp, group_by(cfecnew, p_holder, year) %>% 
    summarize(totRev = sum(g_earn)))
temp$totRev.spec = temp$totRev.spec / temp$totRev

mkt = group_by(cfecnew, p_fshy, year, specn) %>% 
  summarize(totRev.mkt.spec = sum(g_earn)) 
mkt = left_join(mkt, group_by(mkt, p_fshy, year) %>% 
  summarize(totRev.mkt = sum(totRev.mkt.spec))) 
mkt$totRev.mkt.spec = mkt$totRev.mkt.spec / mkt$totRev.mkt
names(mkt)[1] = "permit"

# combine individual and market datasets
temp = left_join(temp, mkt)

# Calculate squared proportions
temp$p2 = (temp$totRev.spec - temp$totRev.mkt.spec)^2

cfecAnnual = left_join(cfecAnnual, dplyr::select(temp, p_holder, year, p2))

@@

cfecAnnual$permit.prev = cfecAnnual$permit[idx]
cfecAnnual$ngear.prev = cfecAnnual$ngear[idx]
cfecAnnual$nspec.prev = cfecAnnual$nspec[idx]
cfecAnnual$npermit.prev = cfecAnnual$npermit[idx]
cfecAnnual$revenue.prev = cfecAnnual$revenue[idx]
cfecAnnual$specdiv.prev = cfecAnnual$specDiv[idx]
cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
cfecAnnual$p2.prev = cfecAnnual$p2[idx]

@@

cfecAnnual.salmon = cfecAnnual[cfecAnnual$salmon=="Y",]

cfecAnnual.salmon = cfecAnnual.salmon[is.na(cfecAnnual.salmon$specDiv.change)==F, ]
saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
```

>>>>

cfecAnnual = group_by(cfecnew, p_holder, year) %>%
  summarize(permit = p_fshy[1], boat = cadfg2[1], 
    days = diff(range(day) + 1), 
    ngear = length(unique(gearn)),
    nspec = length(unique(specn)),
    npermit = length(unique(p_fshy)),
    revenue = sum(g_earn),
    length = mean(VLength, na.rm=T))

cfecAnnual = group_by(cfecAnnual, p_holder, year) %>%
  filter(npermit == 1)

@@

#temp = group_by(cfecnew, p_holder, year, specn) %>% 
#  summarize(totRev.spec = sum(g_earn), 
#    permit = p_fshy[1])
#temp = left_join(temp, group_by(cfecnew, p_holder, year) %>% 
#    summarize(totRev = sum(g_earn)))
#temp$totRev.spec = temp$totRev.spec / temp$totRev

#mkt = group_by(cfecnew, p_fshy, year, specn) %>% 
#  summarize(totRev.mkt.spec = sum(g_earn)) 
#mkt = left_join(mkt, group_by(mkt, p_fshy, year) %>% 
#  summarize(totRev.mkt = sum(totRev.mkt.spec))) 
#mkt$totRev.mkt.spec = mkt$totRev.mkt.spec / mkt$totRev.mkt
#names(mkt)[1] = "permit"

# combine individual and market datasets
#temp = left_join(temp, mkt)

# Calculate squared proportions
#temp$p2 = (temp$totRev.spec - temp$totRev.mkt.spec)^2

#cfecAnnual = left_join(cfecAnnual, dplyr::select(temp, p_holder, year, p2))

@@

cfecAnnual$permit.prev = cfecAnnual$permit[idx]
cfecAnnual$ngear.prev = cfecAnnual$ngear[idx]
cfecAnnual$nspec.prev = cfecAnnual$nspec[idx]
cfecAnnual$npermit.prev = cfecAnnual$npermit[idx]
cfecAnnual$revenue.prev = cfecAnnual$revenue[idx]
cfecAnnual$specdiv.prev = cfecAnnual$specDiv[idx]
cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
#cfecAnnual$p2.prev = cfecAnnual$p2[idx]

@@

cfecAnnual.salmon = cfecAnnual[cfecAnnual$salmon=="Y",]
cfecAnnual.nonsalmon = cfecAnnual[cfecAnnual$salmon=="N",]

cfecAnnual.salmon = cfecAnnual.salmon[is.na(cfecAnnual.salmon$specDiv.change)==F, ]
cfecAnnual.nonsalmon = cfecAnnual.nonsalmon[is.na(cfecAnnual.nonsalmon$specDiv.change)==F, ]

saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
saveRDS(cfecAnnual.nonsalmon, file="salmon_linearModeling_complete.rds")
```

We can look at whether there's clear combinations of permit types that are combinations -- most of these people either have a single local permit (PWS, Bristol Bay, etc) or have multiple statewide permits.

cfecnew$area = substr(cfecnew$p_fshy,4,4)
cfecnew$area[cfecnew$area %in% c("A","C")] = "Southeast"
cfecnew$area[cfecnew$area == "B"] = "Statewide"
cfecnew$area[cfecnew$area == "D"] = "Yakutat"
cfecnew$area[cfecnew$area == "E"] = "PWS"
cfecnew$area[cfecnew$area == "H"] = "CookInlet"
cfecnew$area[cfecnew$area == "J"] = "Westward"
cfecnew$area[cfecnew$area == "K"] = "Kodiak"
cfecnew$area[cfecnew$area == "L"] = "Chignik"
cfecnew$area[cfecnew$area == "M"] = "AlaskaPeninsula"
cfecnew$area[cfecnew$area == "N"] = "Nelson"
cfecnew$area[cfecnew$area == "Q"] = "BeringSea"
cfecnew$area[cfecnew$area == "T"] = "BristolBay"
cfecnew$area[cfecnew$area == "X"] = "Kotzebue"
cfecnew$area[cfecnew$area == "Z"] = "NortonSound"

cfecnew$spec = substr(cfecnew$p_fshy,1,1)
cfecnew$spec[cfecnew$spec == "B"] = "Halibut"
cfecnew$spec[cfecnew$spec == "C"] = "Sablefish"
cfecnew$spec[cfecnew$spec %in% c("D","E","K","T")] = "Crab"
cfecnew$spec[cfecnew$spec %in% c("F")] = "Freshwater"
cfecnew$spec[cfecnew$spec %in% c("G","H","L")] = "Herring"
cfecnew$spec[cfecnew$spec %in% c("I")] = "Lingcod"
cfecnew$spec[cfecnew$spec %in% c("A","J","N","R","W")] = "Shellfish"
cfecnew$spec[cfecnew$spec %in% c("O","P","Q","U","Z")] = "Invert"
cfecnew$spec[cfecnew$spec %in% c("M")] = "Misc"
cfecnew$spec[cfecnew$spec %in% c("S")] = "Salmon"
cfecnew$area.spec = paste(cfecnew$area, cfecnew$spec, sep="_")

cfecAnnual.multiplePermits = group_by(cfecnew, p_holder, year) %>%
  summarize(npermit = length(unique(p_fshy)), 
    allpermits = paste(sort(unique(p_fshy)), collapse = " "),
    areaspec = paste(sort(unique(area.spec)), collapse = " ")) %>%
  filter(npermit > 1)

```

## Model fitting

Start by fitting some lmer models to revenue for the non-salmon people as an example (smaller dataset).

```{r}
library(lme4)

cfecAnnual.salmon$spec = substr(cfecAnnual.salmon$permit,1,1)
cfecAnnual.salmon$area = substr(cfecAnnual.salmon$permit,4,4)

full.mod = lmer(log(revenue) ~ I(log(revenue.prev)) + specDiv +specDiv.change+ (1+specDiv|permit) + length + days, data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])

full.mod.ind = lmer(log(revenue) ~ permit*as.factor(year) + specDiv + (1+specDiv|permit) + length + days + (1|p_holder), data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])

group_by(cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),], p_holder) %>% 
  summarize(earn = mean(log(revenue)), cv = sd(log(revenue)) / earn, div = mean(specDiv), permit = permit[1]) %>%
  as.data.frame %>%
  group_by(permit) %>% 
  mutate(n = n()) %>% 
  filter(n > 50) %>% 
  ggplot(aes(x = div, y = earn, color = cv)) + geom_point() + facet_wrap(~ permit, scales = "free")
```

### 2016-06-14 portfolio/analysis/4a-createData_linear_modeling.Rmd
Start rearranging the files

saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
saveRDS(cfecAnnual.nonsalmon, file="salmon_linearModeling_complete.rds")
```

We can look at whether there's clear combinations of permit types that are combinations -- most of these people either have a single local permit (PWS, Bristol Bay, etc) or have multiple statewide permits.

@@

## Model fitting

Start by fitting some lmer models to revenue for the non-salmon people as an example (smaller dataset).

```{r}
library(lme4)

cfecAnnual.salmon$spec = substr(cfecAnnual.salmon$permit,1,1)
cfecAnnual.salmon$area = substr(cfecAnnual.salmon$permit,4,4)

full.mod = lmer(log(revenue) ~ I(log(revenue.prev)) + specDiv +specDiv.change+ (1+specDiv|permit) + length + days, data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])

full.mod.ind = lmer(log(revenue) ~ permit*as.factor(year) + specDiv + (1+specDiv|permit) + length + days + (1|p_holder), data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])

group_by(cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),], p_holder) %>% 
  summarize(earn = mean(log(revenue)), cv = sd(log(revenue)) / earn, div = mean(specDiv), permit = permit[1]) %>%
  as.data.frame %>%
  group_by(permit) %>% 
  mutate(n = n()) %>% 
  filter(n > 50) %>% 
  ggplot(aes(x = div, y = earn, color = cv)) + geom_point() + facet_wrap(~ permit, scales = "free")
```

>>>>

saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
saveRDS(cfecAnnual.nonsalmon, file="nonsalmon_linearModeling_complete.rds")

## Code for looking at permit 
We can look at whether there's clear combinations of permit types that are combinations -- most of these people either have a single local permit (PWS, Bristol Bay, etc) or have multiple statewide permits.

@@
